---
title: "covid 19"
author: "Esteffani Maldonado"
date: "9/30/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
    
Based on the first bar chart, we can see that there is a higher number of females in this study. However, the amount of males that died is actually higher than the amount of females. It would be expected that if it were indiscriminate, we would see the same amount of deaths whether or not you were male or female.
  



  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The first line of the first table tells us that of all the people that remain alive, 56% of them are females and 44% are males. The second line of the first tells us that of all of the people who are dead, 36% are females and 64% are males.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Looking at the chi squared test, there were about 400 more females than there were males. When we look at the number of deaths, 25 females died and 44 males died. Almost twice as many males died than females even though there were less males in the study to begin with. Looking at the fisher test, the ratio of males dead to females dead is about 2.3 to 1, meaning the odds of dying for men is more than two times greater than the odds of dying for women.

## Conclusions

We used a categorical-categorical helper file to solve this problem. We observed two barcharts as well as a chi squared test and a fisher test. From the bar charts, we were able to see that more men had died from covid than women. From the chi squared test, we saw that the observed counts had not lined up with the expected counts. It was expected that more women would die than men because there was a larger number of women in this sample, but instead more men had actually died than women. This meant that to answer the question "Is dying independent of sex?" we would have to answer that no, men are more likely to die than women.
  
